---
title: "Dot Plot Brief" 
subtitle: ""
author: "Christian Hower"
date: '2018-06-20'
output: 
  html_document:
    theme: spacelab
    highlight: monochrome
    toc: false
    toc_float: true
    toc_depth: 4
    #code_folding: hide
    css: styles.css
--- 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE ,warning=FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(gapminder)
```


### <span style="color:black">How to Read it and What to look for</span>
The dot plot is a type of graph used to display the distribution of a continuous variable, often grouping by category. To read a dot plot , the reader should look for how mass is distributed along the plot. This indicates the frequency of observations at the corresponding value.   


### <span style="color:black">Example:</span>

```{r, echo = FALSE, message = FALSE, warning = FALSE}
gap_country <- gapminder %>% filter(year == "2007") %>%  group_by(continent) %>% summarize(StdDev = sd(gdpPercap), Min = min(gdpPercap),Mean = mean(gdpPercap), Max = max(gdpPercap))
```



```{r, warning = FALSE, message = FALSE} 
#Spread of countries gdp by continent
gap_country <- gapminder %>% group_by(continent, country, year) %>% filter(year == "2007")

ggplot(gap_country,aes(x = continent, y = gdpPercap,  fill = continent ))+
  geom_dotplot(binwidth = .01, aes(fill = continent))+
  geom_dotplot(binaxis = "y", stackdir = "center")+
  labs(x = "Continent", y = "GDP Per Capita",
       title = "Distribution of GDP Per Capita in 2007 by Continent",
       caption = "Many countries are still behind in per capita income.")+
  scale_fill_discrete(name = "")
```


### <span style="color:black">Data</span>
This example uses the gapminder dataset which details global gross domestic product. The dataset contains measures of gdp by country and continent over time. This dataset can be found at [gapminder.org]("https://www.gapminder.org/data/")

### <span style="color:black">Presentation Tips</span>

* **Interactivity:** One interactive feature that could be useful is to display information about a point while hovering over it.

* **Annotation:** Annotation can be used to identify outliers or other points of interest.

* **Color:** Color can be used to represent levels of a category.

* **Composition:** Vertical or horizontal placement of the plots may be easier to compare depending on context.
 
